The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2024-08-28 ms/ms.tex
more edits

\begin{keypoints}
\item enter point 1 here
\end{keypoints}

\begin{abstract}
\begin{itemize}
\item 211 parameters perturbed across six forcing scenarios
\item expansive dataset available which can be used to identify the most influential parameters on a wide range of output variables globally, by biome, or by plant functional type
\item parameter effects can exceed scenario effects
\item small number of parameters explains a large fraction of variance 
\item most important parameters can vary regionally and also based on climate forcing
\item software infrastructure developed to facilitate routine investigation of parameter sensitivity and uncertainty, as well as automated calibration
\end{itemize}
\end{abstract}

\section{Introduction}
Water availability, land temperature extremes, fire risk, and crop productivity will all see impacts from climate change, and are among the many processes represented within the terrestrial components of Earth System Models. Understanding how these processes respond to and influence CO$_2$ concentrations is a critical facet of climate change research. Uncertainty in climate model projections varies by domain and generally increases with extended time horizons \cite{koven2022}. Land processes substantially influence climate directly through, for example, evapotranspiration \cite{zarakas2024}, and indirectly through carbon-climate feedbacks \cite{cox2000}.
Making centennial-scale projections of the cumulative terrestrial carbon sink gas been especially challenging, with high uncertainty persisting across model generations \cite{friedlingstein2014,arora2020}, and limited efficacy of emergent constraints. A portion of this uncertainty is irreducible, inherent to the challenge of predicting vegetation dynamics in a novel climate \cite{lovenduski2017}. Still, our expectation is that some of this uncertainty is indeed reducible, if we can more effectively utilize the ongoing expansion of observational data sources from remote sensing, meteorological stations, flux towers, and field campaigns \cite{worden2021,baldocchi2024}. The capability to efficiently ingest these data and improve simulation performance is especially valuable for actionable science \cite{cheng2023}. However, given increasingly comprehensive land modeling systems, and the wide array of observational products, several technical hurdles exist that hinder effective model development and calibration.

@@

The process of parameter estimation for the complex land models typically embedded within Earth System Models is challenging on account of the intrinsic complexity of the heterogeneous land surface, the diversity and plasticity of plant and microbial life, and the multiplicity of domains that impact the terrestrial biosphere (e.g, hydrology, snow and ice sheets, biogeochemistry, physiology, land management, etc.). The number of model parameters required to represent all the relevant processes is large, and model simulations are relatively expensive, presenting a major barrier to robust objective calibration, model interpretation, and development \cite{fisher2020,dagon2020}.
Thorough assessment of the parametric sensitivity of land models in a manner that is repeatable, open, and integrated with ongoing code development is desirable but challenging \cite{hourdin2017,balaji2022}. It typically requires extensive computational time,  software engineering, and domain-specific expert scientific knowledge.
Our work here tries to alleviate each of these challenges in turn, at least within the context of one land model.
While we are not yet at the point of automated calibration, we present herein several innovations that will serve to make an automated calibration system more technically feasible.

>>>>

\begin{keypoints}
\item We constructed a large parameter pertubration ensemble of the Community Land Model, version 5.1, perturbing 211 parameters independently across six forcing scenarios.
\item Parameter effects can exceed scenario effects and parameter rankings vary by biome and based on the forcing scenario.
\item The software infrastructure developed for this experiment has greatly reduced the human and computer time needed for constructing this type of ensemble, which can facilitate more routine PPEs.
\end{keypoints}

\begin{abstract}
Comprehensive land models are subject to significant parametric uncertainty, which can be hard to quantify due to the large number of parameters and high model computational costs.
We constructed a large parameter pertubration ensemble (PPE) for the Community Terrestrial Systems Model in the Community Land Model with biogeochemistry configuration (CLM-BGC).
We performed more than 2000 simulations perturbing 211 parameters across six forcing scenarios.
This provides an expansive dataset, which can be used to identify the most influential parameters on a wide range of output variables globally, by biome, or by plant functional type.
We found that parameter effects can exceed scenario effects and that a small number of parameters explains a large fraction of variance across our ensemble. 
The most important parameters can vary regionally and also based on climate forcing.
The software infrastructure developed for this experiment has greatly reduced the human and computer time needed for CLM PPEs, which can facilitate routine investigation of parameter sensitivity and uncertainty, as well as automated calibration.
\end{abstract}

\section{Introduction}
Water availability, land temperature extremes, fire risk, and crop productivity will all see impacts from climate change, and are among the many processes represented within the terrestrial components of Earth System Models. Understanding how these processes respond to rising CO$_2$ and even influence CO$_2$ concentration trajectories is a critical facet of climate change research. Land processes substantially influence climate directly through, for example, evapotranspiration \cite{zarakas2024}, and indirectly through carbon-climate feedbacks \cite{cox2000}.
Uncertainty in climate model projections varies by domain and generally increases with extended time horizons \cite{koven2022}. 
Making centennial-scale projections of the cumulative terrestrial carbon sink gas been especially challenging, with high uncertainty persisting across model generations \cite{friedlingstein2014,arora2020}, and limited efficacy of emergent constraints \cite{liu2024}. A portion of this uncertainty is irreducible, inherent to the challenge of predicting vegetation dynamics in a novel climate \cite{lovenduski2017}. Still, our expectation is that some of this uncertainty is indeed reducible, if we can more effectively utilize the ongoing expansion of observational data sources from remote sensing, meteorological stations, flux towers, and field campaigns \cite{worden2021,baldocchi2024}. The capability to efficiently ingest these data and improve simulation performance is especially valuable for actionable science \cite{cheng2023}. However, given increasingly comprehensive land modeling systems, and the wide array of observational products, several technical hurdles exist that hinder effective model development and calibration.

@@

The process of parameter estimation for the complex land models typically embedded within Earth System Models is challenging on account of the intrinsic complexity of the heterogeneous land surface, the diversity and plasticity of plant and microbial life, and the multiplicity of domains that comprise the terrestrial biosphere (e.g, hydrology, snow and ice sheets, biogeochemistry, physiology, land management, etc.). The number of model parameters required to represent all the relevant processes is large, and model simulations are relatively expensive, presenting a major barrier to robust objective calibration, model interpretation, and development \cite{fisher2020,dagon2020}.
Thorough assessment of the parametric sensitivity of land models in a manner that is repeatable, open, and integrated with ongoing code development is desirable but challenging \cite{hourdin2017,balaji2022}. It typically requires extensive computational time,  software engineering, and domain-specific expert scientific knowledge.
Our work here tries to alleviate each of these challenges in turn, at least within the context of one land model.
While we are not yet at the point of automated calibration, we present herein several innovations that will serve to make an automated calibration system more technically feasible.
